---
title: "figures"
author: "Horst-Heinen"
date: "29/09/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

# COVARIATES - FIGURES

# IMPROVING DATA COMPATIBILITY ON SPECTRAL LIBRARIES: SOIL ORGANIC CARBON PEDOTRANSFER FUNCTIONS FOR SOUTHERN BRAZIL

## Objetive

Experimental procedure to harmonize SOC data (DC, WCt, and WCc) from the southern Brazilian spectral library considering its geographic, topography, pedologic, and spectra data.


```{r}
library(raster)
library(sp)
library(tidyr)
library(magrittr)
library(factoextra)
library(ggplot2)
library(gridExtra)
library(grid)
library(gridGraphics)
library(latticeExtra)
library(lattice)
library(dplyr)
library(glue)
```


### Soil data
The dataset is stored in the folder `data`.

```{r}
data <- read.csv("data/data.csv", sep = ";")
```

### Figure 1

Empirical probability density of carbon (A, C, and E) acontent in soil samples  according to the three analytical methods and the theoretical normal probability density function (dashed line).
Key characteristics of the soil samples and its site, such as the clay content (B), land use/land cover (D), and parental material (F).

```{r, eval = FALSE}
l <- layer(
  panel.mathdensity(
    dmath = dnorm, args = list(mean = mean(x, na.rm = TRUE), sd = sd(x, na.rm = TRUE)), col = 'black', lty = 'dashed'))

p1 <- histogram(
  ~ DC, data, 
  xlab = expression('SOC, Dry combustion [%]'), col = 'lightgray', type = "percent", 
  panel = function (...) {
    panel.grid(v = -1, h = -1)
    panel.histogram(...)
    lattice::panel.rug(...)
  },
  page = function (n) {
    grid::grid.text(label = "a)", x = grid::unit(0.04, "npc"), y = grid::unit(0.95, "npc"))
  }) + l

p2 <- histogram(
  ~ WCt, data, 
  xlab = expression('SOC,  Wet digestion + colorimetry [%]'), col = 'lightgray', type = "percent", 
  panel = function (...) {
    panel.grid(v = -1, h = -1)
    panel.histogram(...)
    lattice::panel.rug(...)
  },
  page = function (n) {
    grid::grid.text(label = "b)", x = grid::unit(0.04, "npc"), y = grid::unit(0.95, "npc"))
  }) + l
  

p3 <- histogram(
  ~ WCc, data, 
  xlab = expression('SOC, Wet digestion + titration [%]'), col = 'lightgray', type = "percent", 
  panel = function (...) {
    panel.grid(v = -1, h = -1)
    panel.histogram(...)
    lattice::panel.rug(...)
  },
  page = function (n) {
    grid::grid.text(label = "c)", x = grid::unit(0.04, "npc"), y = grid::unit(0.95, "npc"))
  }) + l


png("result/chap03-hist-soc.png")
gridExtra::grid.arrange(p1, p2, p3,  ncol = 2)
dev.off()
```

### Covariates

#### Spectrais

```{r}
specMinAbund <- data[33:35]
specOrg <- data[36:59]
```

## ORGANIC FEATURES
```{r}

# plotara curva espectral e indicar as faixar das feições orgânicas
```

### Figure - HISTOGRAM MINERAL COMPOSITION

Empirical probability density of spectra features in soil samples  according to the three analytical methods and the theoretical normal probability density function (dashed line).

```{r, eval = FALSE}
p1 <- histogram(
  ~ rKao, data, 
  xlab = expression('Relative abundance of Kaolinite'), type = "percent", 
  col = 'lightgray',
  panel = function (...) {
    lattice::panel.grid(v = -1, h = -1)
    lattice::panel.histogram(...)
    lattice::panel.rug(...)
  },
  page = function (n) {
    grid::grid.text(label = "b)", x = grid::unit(0.04, "npc"), y = grid::unit(0.95, "npc"))
  }) +
  latticeExtra::layer(panel.abline(v = c(350, 600), lty = 'dotted'))

p2 <- histogram(
  ~ rGoe, data, 
  xlab = expression('Relative abundance of Goethite'), type = "percent", 
  col = 'lightgray',
  panel = function (...) {
    lattice::panel.grid(v = -1, h = -1)
    lattice::panel.histogram(...)
    lattice::panel.rug(...)
  },
  page = function (n) {
    grid::grid.text(label = "c)", x = grid::unit(0.04, "npc"), y = grid::unit(0.95, "npc"))
  }) +
  latticeExtra::layer(panel.abline(v = c(350, 600), lty = 'dotted'))
  

p3 <- histogram(
  ~ rHem, data, 
  xlab = expression('Relative abundance of Hematite'), ylab = 'Percent of total', 
  col = 'lightgray',
  panel = function (...) {
    lattice::panel.grid(v = -1, h = -1)
    lattice::panel.histogram(...)
    lattice::panel.rug(...)
  },
  page = function (n) {
    grid::grid.text(label = "d)", x = grid::unit(0.04, "npc"), y = grid::unit(0.95, "npc"))
  }) +
  latticeExtra::layer(panel.abline(v = c(350, 600), lty = 'dotted'))


p4 <- histogram(
  ~ clay, data,
  xlab = expression('Clay content, g kg'^'-1'), ylab = 'Percent of total', 
  col = 'lightgray',
  panel = function (...) {
    lattice::panel.grid(v = -1, h = -1)
    lattice::panel.histogram(...)
    lattice::panel.rug(...)
  },
  page = function (n) {
    grid::grid.text(label = "a)", x = grid::unit(0.04, "npc"), y = grid::unit(0.95, "npc"))
  }) +
  latticeExtra::layer(panel.abline(v = c(350, 600), lty = 'dotted'))

png("result/hist-relative-mineras.png")
gridExtra::grid.arrange(p4, p1, p2, p3, ncol = 2)
dev.off()
```

```{r}
data$Klevel <- cut(data$rKao, 
                   breaks=c(-Inf, 0.7, 1.4, Inf), 
                   labels=c("Klow","Kmed","Khigh"))

data$Glevel <- cut(data$rGoe, 
                   breaks=c(-Inf, 0.4, 0.8, Inf), 
                   labels=c("Glow","Gmed","Ghigh"))

data$Hlevel <- cut(data$rHem, 
                   breaks=c(-Inf, 0.4, 0.8, Inf), 
                   labels=c("Hlow","Hmed","Hhigh"))
```


### Figure 2

Correlation chart

```{r}
png("result/corr-soc-px.png", width = 480 * 4, height = 480 * 4, res = 72 * 3)
data %>% 
  dplyr::select(DC, WCc, WCt, rKao, rGoe, rHem) %>%
  psych::pairs.panels(method = "spearman", 
                      bg=c("red","yellow","blue")[as.factor(data$local)],)

dev.off()

```

#### Coordinates

```{r}
wgs84utm22s <- sp::CRS('+proj=utm +zone=22 +south +ellps=WGS84 +datum=WGS84 +units=m +no_defs')
sirgas2000 <- sp::CRS('+proj=longlat +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +no_defs')
#sp::spTransform(wgs84utm22s)
coordinates(data) <- ~ coord_x + coord_y
proj4string(data) <- CRS("+proj=longlat +datum=WGS84")
```

##### topogáfica

```{r}
p1 <- histogram(
  ~ twi, data, 
  xlab = expression('Topographic wetness index'), type = "percent", 
  col = 'lightgray',
  panel = function (...) {
    lattice::panel.grid(v = -1, h = -1)
    lattice::panel.histogram(...)
    lattice::panel.rug(...)
  },
  page = function (n) {
    grid::grid.text(label = " ", x = grid::unit(0.04, "npc"), y = grid::unit(0.95, "npc"))
  }) +
  latticeExtra::layer(panel.abline(v = c(350, 600), lty = 'dotted'))

png("result/chap03-hist-twi.png")
gridExtra::grid.arrange(p1, ncol = 1)
dev.off()
```


### Land use and land cover (LULC)
https://code.earthengine.google.com/?accept_repo=users/mapbiomas/user-toolkit 



`land-use`
classified using FAO guidelines for soil description. The following codes are used:

* FS (Floresta): Semi-deciduous forest (vegetation slightly disturbed)
* U (Vegetação secundária): Not used and not managed, vegetation strongly disturbed by clearing, burning,
  ploughing (secondary vegetation, mix of semi-deciduous shrubs and tall grassland)
* AA (Agricultura): Annual field cropping
* FP (Silvicultura): Plantation forestry
* HE (Campo nativo): Animal husbandry (extensive grazing)

```{r}
#MAPBIOMAS
data$lulc <-  as.factor(data$lulc)

p1 <- barchart(
  data$lulc, xlab = 'Land cover', ylab = 'Percent of total', horizontal = FALSE, 
  col = 'lightgray',
  panel = function (...) {
    lattice::panel.grid(h = -1, v = 0)
    lattice::panel.barchart(...)
  },
  page = function (n) {
    grid::grid.text(label = "a)", x = grid::unit(0.04, "npc"), y = grid::unit(0.95, "npc"))
  })

p2 <- histogram(
  ~ twi, data, 
  xlab = expression('Topographic wetness index'), type = "percent", 
  col = 'lightgray',
  panel = function (...) {
    lattice::panel.grid(v = -1, h = -1)
    lattice::panel.histogram(...)
    lattice::panel.rug(...)
  },
  page = function (n) {
    grid::grid.text(label = "b)", x = grid::unit(0.04, "npc"), y = grid::unit(0.95, "npc"))
  }) +
  latticeExtra::layer(panel.abline(v = c(350, 600), lty = 'dotted'))

png("result/chap03-hist-lulc-twi.png")
gridExtra::grid.arrange(p1, p2, ncol = 2)
dev.off()
```


### Climate

WorldClim
```{r}
l = list.files(path = 'D:/Artigos/paper2/data/raster/climate/select', pattern = glob2rx('*.tif'), full.names = TRUE)

l <- lapply(l, raster)
lim = raster::stack(l)
proj4string(lim) <- CRS("+proj=longlat +datum=WGS84")
```

Empirical probability density of environmental covariates (A, C, and E) acontent in soil samples  according to the three analytical methods and the theoretical normal probability density function (dashed line).
Key characteristics of the soil samples and its site, such as the clay content (B), land use/land cover (D), and parental material (F).

```{r, eval = FALSE}

p1 <- histogram(
  ~ bio1c2, data, xlab = expression('Annual mean temperature (ºC)'), ylab = 'Percent of total', 
  col = 'lightgray',
  panel = function (...) {
    lattice::panel.grid(v = -1, h = -1)
    lattice::panel.histogram(...)
    lattice::panel.rug(...)
  },
  page = function (n) {
    grid::grid.text(label = "a)", x = grid::unit(0.04, "npc"), y = grid::unit(0.95, "npc"))
  }) +
  latticeExtra::layer(panel.abline(v = c(350, 600), lty = 'dotted'))


p2 <- histogram(
  ~ bio3c2, data, xlab = expression('Isothermality'), ylab = 'Percent of total', 
  col = 'lightgray',
  panel = function (...) {
    lattice::panel.grid(v = -1, h = -1)
    lattice::panel.histogram(...)
    lattice::panel.rug(...)
  },
  page = function (n) {
    grid::grid.text(label = "b)", x = grid::unit(0.04, "npc"), y = grid::unit(0.95, "npc"))
  }) +
  latticeExtra::layer(panel.abline(v = c(350, 600), lty = 'dotted'))

p3 <- histogram(
  ~ bio7c2, data, xlab = expression('Temperature annual range (ºC)'), ylab = 'Percent of total', 
  col = 'lightgray',
  panel = function (...) {
    lattice::panel.grid(v = -1, h = -1)
    lattice::panel.histogram(...)
    lattice::panel.rug(...)
  },
  page = function (n) {
    grid::grid.text(label = "c)", x = grid::unit(0.04, "npc"), y = grid::unit(0.95, "npc"))
  }) +
  latticeExtra::layer(panel.abline(v = c(350, 600), lty = 'dotted'))


p4 <- histogram(
  ~ bio12c2, data, 
  xlab = expression('Annual precipitation (mm)'), col = 'lightgray', type = "percent", 
  panel = function (...) {
    panel.grid(v = -1, h = -1)
    panel.histogram(...)
    lattice::panel.rug(...)
  },
  page = function (n) {
    grid::grid.text(label = "d)", x = grid::unit(0.04, "npc"), y = grid::unit(0.95, "npc"))
  }) + l
  
p5 <- histogram(
  ~ bio15c2, data, 
  xlab = expression('Precipitation Seasonality'), col = 'lightgray', type = "percent", 
  panel = function (...) {
    panel.grid(v = -1, h = -1)
    panel.histogram(...)
    lattice::panel.rug(...)
  },
  page = function (n) {
    grid::grid.text(label = "e)", x = grid::unit(0.04, "npc"), y = grid::unit(0.95, "npc"))
  }) + l




png("result/chap03-hist-climate22.png")
gridExtra::grid.arrange(p1, p2, p3, p4, p5, ncol = 2)
dev.off()
```


classified using European Soil Map (HYPRES) guidelines for soil description. The following codes are used:
http://www.macaulay.ac.uk/hypres/hypressoil.html.

* VF = Very fine (clay >= 60)
* FN = Fine (<600 clay >350)
* ME = Medium (clay <350)

```{r}
data$texture <- cut(data$clay, 
                   breaks=c(-Inf, 350, 600, Inf), 
                   labels=c("ME","FN","VF"))
```

I create new categorical predictor variables using data on LULC (forest, agriculture, grazing) and soil texture (medium,fine, very fine). These are dummy variables composed of zeros and ones.

```{r}
data <-
  data %>%
  mutate(
    FS = ifelse(landuse == "FS", 1, 0),
    U =  ifelse(landuse == "U", 1, 0),
    AA = ifelse(landuse == "AA", 1, 0),
    FP = ifelse(landuse == "FP", 1, 0),   
    HE = ifelse(landuse == "HE", 1, 0),
    
    AV = ifelse(geology == "AV", 1, 0),
    BV = ifelse(geology == "BV", 1, 0),
    SE = ifelse(geology == "SE", 1, 0),

    ME = ifelse(texture == "ME", 1, 0),
    FI = ifelse(texture == "FI", 1, 0),
    VF = ifelse(texture == "VF", 1, 0))
```

#ALL COVARS
```{r, eval = FALSE}
p1 <- histogram(
  ~ rKao, data, 
  xlab = expression('Relative abundance of Kaolinite'), type = "percent", 
  col = 'lightgray',
  panel = function (...) {
    lattice::panel.grid(v = -1, h = -1)
    lattice::panel.histogram(...)
    lattice::panel.rug(...)
  },
  page = function (n) {
    grid::grid.text(label = "b)", x = grid::unit(0.04, "npc"), y = grid::unit(0.95, "npc"))
  })

p2 <- histogram(
  ~ rGoe, data, 
  xlab = expression('Relative abundance of Goethite'), type = "percent", 
  col = 'lightgray',
  panel = function (...) {
    lattice::panel.grid(v = -1, h = -1)
    lattice::panel.histogram(...)
    lattice::panel.rug(...)
  },
  page = function (n) {
    grid::grid.text(label = "c)", x = grid::unit(0.04, "npc"), y = grid::unit(0.95, "npc"))
  })
  

p3 <- histogram(
  ~ rHem, data, 
  xlab = expression('Relative abundance of Hematite'), ylab = 'Percent of total', 
  col = 'lightgray',
  panel = function (...) {
    lattice::panel.grid(v = -1, h = -1)
    lattice::panel.histogram(...)
    lattice::panel.rug(...)
  },
  page = function (n) {
    grid::grid.text(label = "d)", x = grid::unit(0.04, "npc"), y = grid::unit(0.95, "npc"))
  })


p4 <- histogram(
  ~ clay, data,
  xlab = expression('Clay content, g kg'^'-1'), ylab = 'Percent of total', 
  col = 'lightgray',
  panel = function (...) {
    lattice::panel.grid(v = -1, h = -1)
    lattice::panel.histogram(...)
    lattice::panel.rug(...)
  },
  page = function (n) {
    grid::grid.text(label = "a)", x = grid::unit(0.04, "npc"), y = grid::unit(0.95, "npc"))
  })

png("result/hist-clay.png")
gridExtra::grid.arrange(p4, p1, p2, p3, ncol = 2)
dev.off()

p5 <- histogram(
  ~ bio1c2, data, xlab = expression('Annual mean temperature (ºC)'), ylab = 'Percent of total', 
  col = 'lightgray',
  panel = function (...) {
    lattice::panel.grid(v = -1, h = -1)
    lattice::panel.histogram(...)
    lattice::panel.rug(...)
  },
  page = function (n) {
    grid::grid.text(label = "e)", x = grid::unit(0.04, "npc"), y = grid::unit(0.95, "npc"))
  })


p6 <- histogram(
  ~ bio3c2, data, xlab = expression('Isothermality'), ylab = 'Percent of total', 
  col = 'lightgray',
  panel = function (...) {
    lattice::panel.grid(v = -1, h = -1)
    lattice::panel.histogram(...)
    lattice::panel.rug(...)
  },
  page = function (n) {
    grid::grid.text(label = "f)", x = grid::unit(0.04, "npc"), y = grid::unit(0.95, "npc"))
  })


p7 <- histogram(
  ~ bio7c2, data, xlab = expression('Temperature annual range (ºC)'), ylab = 'Percent of total', 
  col = 'lightgray',
  panel = function (...) {
    lattice::panel.grid(v = -1, h = -1)
    lattice::panel.histogram(...)
    lattice::panel.rug(...)
  },
  page = function (n) {
    grid::grid.text(label = "g)", x = grid::unit(0.04, "npc"), y = grid::unit(0.95, "npc"))
  })


p8 <- histogram(
  ~ bio12c2, data, 
  xlab = expression('Annual precipitation (mm)'), col = 'lightgray', type = "percent", 
  panel = function (...) {
    panel.grid(v = -1, h = -1)
    panel.histogram(...)
    lattice::panel.rug(...)
  },
  page = function (n) {
    grid::grid.text(label = "h)", x = grid::unit(0.04, "npc"), y = grid::unit(0.95, "npc"))
  })
  
p9 <- histogram(
  ~ bio15c2, data, 
  xlab = expression('Precipitation Seasonality (Coefficient of Variation)'), col = 'lightgray', type = "percent", 
  panel = function (...) {
    panel.grid(v = -1, h = -1)
    panel.histogram(...)
    lattice::panel.rug(...)
  },
  page = function (n) {
    grid::grid.text(label = "i)", x = grid::unit(0.04, "npc"), y = grid::unit(0.95, "npc"))
  })

png("result/hist-climate.png")
gridExtra::grid.arrange(p5, p6, p7, p8, p9, ncol = 2)
dev.off()


data$lulc <- as.factor(data$lulc_mapbiomas)
p10 <- barchart(
  data$lulc_mapbiomas, xlab = 'Land cover', ylab = 'Percent of total', horizontal = FALSE, 
  col = 'lightgray',
  panel = function (...) {
    lattice::panel.grid(h = -1, v = 0)
    lattice::panel.barchart(...)
  },
  page = function (n) {
    grid::grid.text(label = "j)", x = grid::unit(0.04, "npc"), y = grid::unit(0.95, "npc"))
  })

p11 <- histogram(
  ~ twi, data, 
  xlab = expression('Topographic wetness index'), type = "percent", 
  col = 'lightgray',
  panel = function (...) {
    lattice::panel.grid(v = -1, h = -1)
    lattice::panel.histogram(...)
    lattice::panel.rug(...)
  },
  page = function (n) {
    grid::grid.text(label = "k)", x = grid::unit(0.04, "npc"), y = grid::unit(0.95, "npc"))
  })

png("result/hist-terrain.png")
gridExtra::grid.arrange(p10, p11, ncol = 2)
dev.off()
```


### Figure spectra

```{r}
data <- read.csv("data/data_soc_methods.csv", sep = ";")
spectra <- data[,c(27:length(data))]
signals <- read.csv("data/signals.csv", sep = ";")
signalsmin <- read.csv("data/signals_min.csv", sep = ";")

#####################################################
ContRem <- prospectr::continuumRemoval(X = spectra, type = "R")
CRR <- ContRem[,c(-1,-ncol(ContRem))]
colnames(CRR)[1:2144] <- paste0("X", c(355:2498))

#########################cálculo ds curvas médias e sd 
####################
geral.spectra <- CRR

mean.geral <- apply(geral.spectra, 2 , FUN = mean) ##### mean 
sd.geral   <- apply(geral.spectra, 2 , FUN = sd) ##### sd
pos.sd.geral        <- -sd.geral ##### pos. and neg. std dev

pos        <- (mean.geral + pos.sd.geral)
neg        <- (mean.geral - pos.sd.geral)


CR.pos.sd.table       <- rbind(mean.geral, pos, neg )

CR.class.mean.pos.neg           <- (data.frame(t(CR.pos.sd.table)))
colnames(CR.class.mean.pos.neg) <- c("Mean","Pos", "Neg")
row.names(CR.class.mean.pos.neg)<- NULL

CR.class.mean.pos.neg$band      <- seq(355,2498)

geralCRR <- CR.class.mean.pos.neg

#####################################################
#########################SPECTRA CURVE
library(ggplot2)
library(grid)
library(gridExtra)

s1 <- ggplot2::ggplot(data = geralCRR) +
  geom_ribbon(aes(ymin = Neg, ymax = Pos, x = band), alpha = 0.4, fill="gray") +
  
  geom_line(aes(x = band, y = Mean), color="black", size = 1, lty =1) +
  geom_vline(xintercept = as.numeric(signals$wav), size = 1, colour = "darkgreen", alpha = .2) +
  
  annotate("rect", xmin = 2078, xmax = 2267, ymin = 0, ymax = 1, fill = "darkgoldenrod", alpha = .2, color = NA) +

  annotate("rect", xmin = 457, xmax = 563, ymin = 0, ymax = 1, fill = "blue", alpha = .2, color = NA) +
  
  annotate("rect", xmin = 776, xmax = 1266, ymin = 0, ymax = 1, fill = "blue", alpha = .2, color = NA) +
  
    annotate("rect", xmin = 455, xmax = 612, ymin = 0, ymax = 1, fill = "brown3", alpha = .2, color = NA) +
  
  annotate("rect", xmin = 765, xmax = 1050, ymin = 0, ymax = 1, fill = "brown3", alpha = .2, color = NA) +


  scale_x_continuous("Wavelength (nm)", limits = c(355,2500)) +
    
  scale_y_continuous("CRR reflectance (nm)") +
  theme_bw() +
   #facet_grid(rows = vars("CRR")) +
    theme(strip.text.y = element_text(size=12, color="black",
                                      face="bold"))


png("result/chap03-spectra.png")
s1
dev.off()
```

