---
title: "SpectraFeatures"
author: "Horst-Heinen"
date: "20/05/2021"
output: html_document
---
# SPECTRA COVARIATES

# IMPROVING DATA COMPATIBILITY ON SPECTRAL LIBRARIES: SOIL ORGANIC CARBON PEDOTRANSFER FUNCTIONS FOR SOUTHERN BRAZIL

## Objetive

Experimental procedure to harmonize SOC data (DC, WCt, and WCc) from the southern Brazilian spectral library considering its geographic, topography, pedologic, and spectra data.

```{r}
#library(devtools)
# install the soilspec package from GitHub
#devtools::install_github("AlexandreWadoux/soilspec", force = FALSE)

myPackages <- c("SDMTools", "tripack", "resemble", "splancs", "sp")
notInstalled <- myPackages[!(myPackages %in% installed.packages()[ , "Package"])]
if(length(notInstalled)) install.packages(notInstalled)


myPackages <- c("asdreader", "RColorBrewer","wavethresh", "MASS",
"pracma", "plyr", "signal", "SDMTools",
"tripack", "resemble", "splancs", "sp",
"scatterplot3d", "prospectr", "RcppArmadillo", "matrixStats",
"clhs", "viridis", "viridisLite", "caret",
"ggplot2", "soiltexture","randomForest","pls",
"Cubist", "lattice", "robustbase", "mvoutlier",
"devtools")
# define which packages are not installed in the current computer
notInstalled <- myPackages[!(myPackages%in%installed.packages()[ , "Package"])]
# install the missing packages
if(length(notInstalled)>0) install.packages(notInstalled)

library(magrittr)
#library(soilspec)
```

# Feature Selection

## Identifying Secondary Clay Minerals and Iron Oxides
Font U.S. Geological Survey digital spectral library

Mineral Diagnostic               wavelength range(s)/nm
----------------------------------------------------------
Kaolinite                         2078–2267
Goethite                          457–563, 776–1266
Hematite                          455–612, 765–1050
-----------------------------------------------------------

#### Functions

```{r trimSpec}
# function for trimming spectra or targeting a specific spectral region of interest

trimSpec <- function(spectra, wavlimits){
datawavs <- as.numeric(colnames(spectra))
# set the limits
limits <- which(datawavs %in% wavlimits)
# mention the index that we keep from the matrix
keptIndex <- seq(limits[1], limits[2], 1)
# keep the index selected previously
trimmedSpectra <- spectra[, keptIndex]
# return the trimmed spectra
keptNames <- datawavs[keptIndex]
colnames(trimmedSpectra) <- keptNames
return(trimmedSpectra)
}
```

```{r HullCRR}
HullCRR <- function (spectra, type = c("R", "A"), wav, ...) 
{
    X <- spectra
    if (type == "A") {
        X <- 1/X
    }
    cHullFun <- function(x, wav) {
        cHull <- sort(chull(c(wav[1] - 1, wav, wav[length(wav)] + 
            1), c(0, x, 0)))
        cHull <- cHull[-c(1, length(cHull))] - 1
        cont <- approx(x = wav[cHull], y = x[cHull], xout = wav, 
            method = "linear")$y
        return(cont)
    }
    cont <- cHullFun(X, wav)
    if (type == "A") {
        hullSpectra <- 1 + X - cont
    }
    else {
        hullSpectra <- X/cont
    }
    if (type == "A") {
        hullSpectra <- 1/hullSpectra
        cont <- 1/cont
        X <- 1/X
    }
    pol <- cbind(wav, as.numeric(hullSpectra[1, ]))
    retval <- list(wave = wav, cHull = hullSpectra, rawSpec = X, 
        continuum = cont, polygon = pol)
    return(retval)
}
```

##### Prepare reference material

```{r imp Ref}
mineralRef <- read.csv("data/mineralRef.csv")
mineralRef <- mineralRef[c("wavelength", "Geothite", "hematite", "kaolinite_114")]
  
# create a sequence of numbers to represent the wavelength
wavelength <- seq(350, 2500, by = 1)
```

```{r kaolinita}
# select reference mineral
ref <- as.data.frame(t(mineralRef$kaolinite_114))
colnames(ref) <- wavelength

lower <- 2078 # diagnostic wavelength range
upper <- 2267 # diagnostic wavelength range

SpecZoom <- trimSpec(ref, 
wavlimits = c(lower, upper)) # spectrum trimming

SpecZoomCR <- HullCRR(spectra = SpecZoom, # continuum removal function
wav = as.numeric(colnames(SpecZoom)),
type = "R")

waveId <- which(SpecZoomCR$cHull==min(SpecZoomCR$cHul))[1] # find minimum

Scr <- SpecZoomCR$rawSpec[waveId] # identify raw reflectance reading
Sl <- SpecZoomCR$continuum[waveId] # calculate band depth

featureDepth <- 1 - (Scr/Sl)
featureArea <- splancs::areapl(SpecZoomCR$polygon) # calculate area of the polygon
featureSlope <- tail(SpecZoomCR$continuum, 1) - SpecZoomCR$continuum[1] # calculate slope of the continuum

kaolin1141summary <- list(name = "kaolinite114",
wave = as.numeric(colnames(SpecZoom[1,])),
CRspectra = SpecZoomCR$cHull,
bandDepth = featureDepth,
featureArea = featureArea,
featureSlope = featureSlope,
continuum = SpecZoomCR$continuum,
contiuumPolygon = SpecZoomCR$polygon,
rawSpectrum = SpecZoomCR$wave)
```

```{r geotita1}
# select reference mineral

ref <- as.data.frame(t(mineralRef$Geothite))
colnames(ref) <- wavelength

lower <- 457 # diagnostic wavelength range
upper <- 563 # diagnostic wavelength range

SpecZoom <- trimSpec(ref, 
wavlimits = c(lower, upper)) # spectrum trimming

SpecZoomCR <- HullCRR(spectra = SpecZoom, # continuum removal function
wav = as.numeric(colnames(SpecZoom)),
type = "R")

waveId <- which(SpecZoomCR$cHull==min(SpecZoomCR$cHul))[1] # find minimum

Scr <- SpecZoomCR$rawSpec[waveId] # identify raw reflectance reading
Sl <- SpecZoomCR$continuum[waveId] # calculate band depth

featureDepth <- 1 - (Scr/Sl)
featureArea <- splancs::areapl(SpecZoomCR$polygon) # calculate area of the polygon
featureSlope <- tail(SpecZoomCR$continuum, 1) - SpecZoomCR$continuum[1] # calculate slope of the continuum

geothita1summary <- list(name = "goethita1",
wave = as.numeric(colnames(SpecZoom[1,])),
CRspectra = SpecZoomCR$cHull,
bandDepth = featureDepth,
featureArea = featureArea,
featureSlope = featureSlope,
continuum = SpecZoomCR$continuum,
contiuumPolygon = SpecZoomCR$polygon,
rawSpectrum = SpecZoomCR$wave)
```

```{r geotita2}
# select reference mineral

spectra <- as.data.frame(t(mineralRef$Geothite))
colnames(spectra) <- wavelength

lower <- 776 # diagnostic wavelength range
upper <- 1266 # diagnostic wavelength range

SpecZoom <- trimSpec(spectra, 
wavlimits = c(lower, upper)) # spectrum trimming

SpecZoomCR <- HullCRR(spectra = SpecZoom, # continuum removal function
wav = as.numeric(colnames(SpecZoom)),
type = "R")

waveId <- which(SpecZoomCR$cHull==min(SpecZoomCR$cHul))[1] # find minimum

Scr <- SpecZoomCR$rawSpec[waveId] # identify raw reflectance reading
Sl <- SpecZoomCR$continuum[waveId] # calculate band depth

featureDepth <- 1 - (Scr/Sl)
featureArea <- splancs::areapl(SpecZoomCR$polygon) # calculate area of the polygon
featureSlope <- tail(SpecZoomCR$continuum, 1) - SpecZoomCR$continuum[1] # calculate slope of the continuum

geothita2summary <- list(name = "goethita2",
wave = as.numeric(colnames(SpecZoom[1,])),
CRspectra = SpecZoomCR$cHull,
bandDepth = featureDepth,
featureArea = featureArea,
featureSlope = featureSlope,
continuum = SpecZoomCR$continuum,
contiuumPolygon = SpecZoomCR$polygon,
rawSpectrum = SpecZoomCR$wave)
```

```{r hematite1}
# select reference mineral

spectra <- as.data.frame(t(mineralRef$hematite))
colnames(spectra) <- wavelength

lower <- 455 # diagnostic wavelength range
upper <- 612 # diagnostic wavelength range

SpecZoom <- trimSpec(spectra, 
wavlimits = c(lower, upper)) # spectrum trimming

SpecZoomCR <- HullCRR(spectra = SpecZoom, # continuum removal function
wav = as.numeric(colnames(SpecZoom)),
type = "R")

waveId <- which(SpecZoomCR$cHull==min(SpecZoomCR$cHul))[1] # find minimum

Scr <- SpecZoomCR$rawSpec[waveId] # identify raw reflectance reading
Sl <- SpecZoomCR$continuum[waveId] # calculate band depth

featureDepth <- 1 - (Scr/Sl)
featureArea <- splancs::areapl(SpecZoomCR$polygon) # calculate area of the polygon
featureSlope <- tail(SpecZoomCR$continuum, 1) - SpecZoomCR$continuum[1] # calculate slope of the continuum

hematite1summary <- list(name = "hematite1",
wave = as.numeric(colnames(SpecZoom[1,])),
CRspectra = SpecZoomCR$cHull,
bandDepth = featureDepth,
featureArea = featureArea,
featureSlope = featureSlope,
continuum = SpecZoomCR$continuum,
contiuumPolygon = SpecZoomCR$polygon,
rawSpectrum = SpecZoomCR$wave)
```

```{r hematite2}
# select reference mineral

spectra <- as.data.frame(t(mineralRef$hematite))
colnames(spectra) <- wavelength

lower <- 765 # diagnostic wavelength range
upper <- 1050 # diagnostic wavelength range

SpecZoom <- trimSpec(spectra, 
wavlimits = c(lower, upper)) # spectrum trimming

SpecZoomCR <- HullCRR(spectra = SpecZoom, # continuum removal function
wav = as.numeric(colnames(SpecZoom)),
type = "R")

waveId <- which(SpecZoomCR$cHull==min(SpecZoomCR$cHul))[1] # find minimum

Scr <- SpecZoomCR$rawSpec[waveId] # identify raw reflectance reading
Sl <- SpecZoomCR$continuum[waveId] # calculate band depth

featureDepth <- 1 - (Scr/Sl)
featureArea <- splancs::areapl(SpecZoomCR$polygon) # calculate area of the polygon
featureSlope <- tail(SpecZoomCR$continuum, 1) - SpecZoomCR$continuum[1] # calculate slope of the continuum

hematite2summary <- list(name = "hematite2",
wave = as.numeric(colnames(SpecZoom[1,])),
CRspectra = SpecZoomCR$cHull,
bandDepth = featureDepth,
featureArea = featureArea,
featureSlope = featureSlope,
continuum = SpecZoomCR$continuum,
contiuumPolygon = SpecZoomCR$polygon,
rawSpectrum = SpecZoomCR$wave)
```


###### QUANTIFY MINERALS

 Comparing Soil Spectra with Spectra of Reference Materials

```{r import data}
data <- read.csv("data/data_covar.csv", sep = ";")
spectra <- data[,c(32:length(data))]

oldNames <- grep("X", names(spectra), value = TRUE)
wavelength <- as.numeric(substring(grep("X", names(spectra), value = T), 2, 20))
colnames(spectra) <- wavelength
```

```{r kao}
AbMineral1 <- data.frame(rKao=numeric())

minSelect <- kaolin1141summary #reference material
lower <- min(minSelect$wave) #region of Interest
upper <- max(minSelect$wave)


for (i in 1:nrow(spectra)) {
 sample <- spectra[i,]
 
      SpecZoom <- trimSpec(sample, wavlimits = c(lower, upper))

      SpecZoomCR <- HullCRR(spectra = SpecZoom, wav = as.numeric(colnames(SpecZoom)), type = "R")
      waveId <- which(SpecZoomCR$cHull==min(SpecZoomCR$cHul))[1] # find wv minimum continuum-removed reflectance
      Scr <- SpecZoomCR$rawSpec[waveId] # identify raw reflectance reading
      Sl <- SpecZoomCR$continuum[waveId] # identify continuum reading
      
  sDepth <- 1 - (Scr/Sl) # calculate band depth
  sArea <- splancs::areapl(SpecZoomCR$polygon)
  sSlope <- tail(SpecZoomCR$continuum, 1) - SpecZoomCR$continuum[1]
  specFit <- as.numeric(cor(t(SpecZoomCR$cHull), t(minSelect$CRspectra))) # correlation
  
  rDepth <- as.numeric(sDepth/minSelect$bandDepth) # relative depth
  rArea <- as.numeric(sArea/minSelect$featureArea) # relative area
  
  FitDepth <- specFit*rDepth # fit x relative depth
  #AbMineral[i,1] <- specFit
  AbMineral1[i,1] <- specFit*rDepth*rArea # fit x depth x area
    }

#View(indexes)
#str(features)
```

```{r goe}
AbMineral2 <- data.frame(rGoe=numeric())

for (i in 1:nrow(spectra)) {
 sample <- spectra[i,]
 
      minSelect1 <- geothita1summary #reference material
      lower1 <- min(minSelect1$wave) #region of Interest
      upper1 <- max(minSelect1$wave)
 
      SpecZoom <- trimSpec(sample, wavlimits = c(lower1, upper1))

      SpecZoomCR <- HullCRR(spectra = SpecZoom, wav = as.numeric(colnames(SpecZoom)), type = "R")
      waveId <- which(SpecZoomCR$cHull==min(SpecZoomCR$cHul))[1] # find wv minimum continuum-removed reflectance
      Scr <- SpecZoomCR$rawSpec[waveId] # identify raw reflectance reading
      Sl <- SpecZoomCR$continuum[waveId] # identify continuum reading
      
  sDepth1 <- 1 - (Scr/Sl) # calculate band depth
  sArea1 <- splancs::areapl(SpecZoomCR$polygon)
  sSlope1 <- tail(SpecZoomCR$continuum, 1) - SpecZoomCR$continuum[1]
  specFit1 <- as.numeric(cor(t(SpecZoomCR$cHull), t(minSelect1$CRspectra))) # correlation
  
###
      minSelect2 <- geothita2summary #reference material
      lower2 <- min(minSelect2$wave) #region of Interest
      upper2 <- max(minSelect2$wave)

      SpecZoom <- trimSpec(sample, wavlimits = c(lower2, upper2))

      SpecZoomCR <- HullCRR(spectra = SpecZoom, wav = as.numeric(colnames(SpecZoom)), type = "R")
      waveId <- which(SpecZoomCR$cHull==min(SpecZoomCR$cHul))[1] # find wv minimum continuum-removed reflectance
      Scr <- SpecZoomCR$rawSpec[waveId] # identify raw reflectance reading
      Sl <- SpecZoomCR$continuum[waveId] # identify continuum reading
      
  sDepth2 <- 1 - (Scr/Sl) # calculate band depth
  sArea2 <- splancs::areapl(SpecZoomCR$polygon)
  sSlope2 <- tail(SpecZoomCR$continuum, 1) - SpecZoomCR$continuum[1]
  specFit2 <- as.numeric(cor(t(SpecZoomCR$cHull), t(minSelect2$CRspectra))) # correlation
  
  ### quantification
  
  rDepth1 <- as.numeric(sDepth1/(sum(minSelect1$bandDepth + minSelect2$bandDepth))) # relative depth
  rArea1 <- as.numeric(sArea1/(sum(minSelect1$featureArea + minSelect2$featureArea))) # relative area
  
  FitDepth1 <- specFit1*rDepth1 # fit x relative depth
  Abund1 <- specFit1*rDepth1*rArea1 # fit x depth x area
  
  rDepth2 <- as.numeric(sDepth2/(sum(minSelect1$bandDepth + minSelect2$bandDepth)))  # relative depth
  rArea2 <- as.numeric(sArea2/(sum(minSelect1$bandDepth + minSelect2$bandDepth)))  # relative area
  
  FitDepth2 <- specFit2*rDepth2 # fit x relative depth
  Abund2 <- specFit2*rDepth2*rArea2 # fit x depth x area
  
  AbMineral2[i,1] <- Abund1 + Abund2 #sum
  }
  
#View(indexes)
#str(features)
```

```{r hem}
AbMineral3 <- data.frame(rHem=numeric())

for (i in 1:nrow(spectra)) {
 sample <- spectra[i,]
 
      minSelect1 <- hematite1summary #reference material
      lower1 <- min(minSelect1$wave) #region of Interest
      upper1 <- max(minSelect1$wave)
 
      SpecZoom <- trimSpec(sample, wavlimits = c(lower1, upper1))

      SpecZoomCR <- HullCRR(spectra = SpecZoom, wav = as.numeric(colnames(SpecZoom)), type = "R")
      waveId <- which(SpecZoomCR$cHull==min(SpecZoomCR$cHul))[1] # find wv minimum continuum-removed reflectance
      Scr <- SpecZoomCR$rawSpec[waveId] # identify raw reflectance reading
      Sl <- SpecZoomCR$continuum[waveId] # identify continuum reading
      
  sDepth1 <- 1 - (Scr/Sl) # calculate band depth
  sArea1 <- splancs::areapl(SpecZoomCR$polygon)
  sSlope1 <- tail(SpecZoomCR$continuum, 1) - SpecZoomCR$continuum[1]
  specFit1 <- as.numeric(cor(t(SpecZoomCR$cHull), t(minSelect1$CRspectra))) # correlation
  
  
###
      minSelect2 <- hematite2summary #reference material
      lower2 <- min(minSelect2$wave) #region of Interest
      upper2 <- max(minSelect2$wave)

      SpecZoom <- trimSpec(sample, wavlimits = c(lower2, upper2))

      SpecZoomCR <- HullCRR(spectra = SpecZoom, wav = as.numeric(colnames(SpecZoom)), type = "R")
      waveId <- which(SpecZoomCR$cHull==min(SpecZoomCR$cHul))[1] # find wv minimum continuum-removed reflectance
      Scr <- SpecZoomCR$rawSpec[waveId] # identify raw reflectance reading
      Sl <- SpecZoomCR$continuum[waveId] # identify continuum reading
      
  sDepth2 <- 1 - (Scr/Sl) # calculate band depth
  sArea2 <- splancs::areapl(SpecZoomCR$polygon)
  sSlope2 <- tail(SpecZoomCR$continuum, 1) - SpecZoomCR$continuum[1]
  specFit2 <- as.numeric(cor(t(SpecZoomCR$cHull), t(minSelect2$CRspectra))) # correlation
  
### quantification
  
  rDepth1 <- as.numeric(sDepth1/(sum(minSelect1$bandDepth + minSelect2$bandDepth))) # relative depth
  rArea1 <- as.numeric(sArea1/(sum(minSelect1$featureArea + minSelect2$featureArea))) # relative area
  
  FitDepth1 <- specFit1*rDepth1 # fit x relative depth
  Abund1 <- specFit1*rDepth1*rArea1 # fit x depth x area
  
  rDepth2 <- as.numeric(sDepth2/(sum(minSelect1$bandDepth + minSelect2$bandDepth)))  # relative depth
  rArea2 <- as.numeric(sArea2/(sum(minSelect1$bandDepth + minSelect2$bandDepth)))  # relative area
  
  FitDepth2 <- specFit2*rDepth2 # fit x relative depth
  Abund2 <- specFit2*rDepth2*rArea2 # fit x depth x area
  
  
  AbMineral3[i,1] <- Abund1 + Abund2 #sum
    }

#View(indexes)
#str(features)
```

```{r cbind}
rAbundMineral <- cbind(AbMineral1, AbMineral2, AbMineral3)
```

## PRONTO
# Demais indices

```{r signals}
dataSignals <- read.csv("data/signals.csv", sep = ";")
signals <- dataSignals[c("X","lower","upper")]
water <- signals[c(10:17),]

signals <- rbind(signals[c(22:nrow(signals)),], water)
```

```{r all}
#features <- data.frame(Depth=numeric(), Area=numeric(), Slope = numeric())
features <- data.frame(Area=numeric())
index <- list()

for (i in 1:nrow(signals)) {
 signal <- signals[i,]
 
    for (j in 1:nrow(spectra)) {
    sample <- spectra[j,]
    
      SpecZoom <- trimSpec(sample, wavlimits = c(signal$lower, signal$upper))

      SpecZoomCR <- HullCRR(spectra = SpecZoom, wav = as.numeric(colnames(SpecZoom)), type = "R")
      waveId <- which(SpecZoomCR$cHull==min(SpecZoomCR$cHul))[1] # find wv minimum continuum-removed reflectance
      Scr <- SpecZoomCR$rawSpec[waveId] # identify raw reflectance reading
      Sl <- SpecZoomCR$continuum[waveId] # identify continuum reading
      
  #features[j,1] <- 1 - (Scr/Sl) # calculate band depth
  features[j,1] <- splancs::areapl(SpecZoomCR$polygon)
  #features[j,3] <- tail(SpecZoomCR$continuum, 1) - SpecZoomCR$continuum[1]
  
  tmp <- features
  it <- i
  #colnames(tmp)[1] <- paste(it, "Depth", sep="")
  colnames(tmp)[1] <- paste(signal$X, it, sep="_")
  #colnames(tmp)[3] <- paste(it, "Slope", sep="")
    }
 
 index[[i]] <- do.call(cbind, tmp)
}
indexs <- do.call(cbind, index)

#View(indexes)
#str(features)
```


# EXPORT
```{r}
index <-cbind(data[1:26], rAbundMineral, indexs)
write.csv(index, "data/data.csv")
```
